Irt plot extra

### 2023-11-07 manuscript_SciRep.Rmd
Update manuscript_SciRep.Rmd

#     institution   : "TUM Department of Sport and Health Sciences (TUM SG), Technical University of Munich, Munich, Germany"

@@

On the other hand, several attempts have been made to quantify received light exposure subjectively with self-report questionnaires (**Supplementary Table 1**), bypassing the cost and intrusiveness issues. However, self-reporting light properties could be challenging for people who lack technical knowledge of light sources. Moreover, individuals may need to consider that, unlike a photometer, the human visual system continuously adapts to brightness [@hurvich_perception_1966], while the signals underlying the non-visual effects of light are independent from perception [@allen_exploiting_2018]. Retrospectively recalling the properties of a light source can further complicate such subjective evaluations. Moreover, measuring light properties alone does not yield any information about how individuals might behave differently regarding diverse light environments such as work, home or outdoors.

To date, little effort has been made to understand and capture these activities. Here, we present the development process of a novel self-reported inventory, the Light Exposure Behaviour Assessment (LEBA), for characterizing diverse light exposure-related behaviours. Notably, we do not claim to be estimating the personal light exposure with these questions. Instead, we aim to assess, in a scalable way, how people behave in relation to light, focusing on habitual patterns that could guide behavioural intervention points.

@@

Exploratory analysis revealed that items of LEBA inventory can be categorized into five major factors: (i) wearing blue light filters; (ii) spending time out doors; (iii) using phone and smartwatch in bed; (iv) using light before bedtime (v) using light in the morning and during daytime. In this stage of analysis, we retained 25 items. the first factor had three items and encapsulated the individual's preference for using blue light filters in different light environments. The second factor contained six items that incorporated the individuals' hours spent outdoors. The third factor contained five items that looked into specific behaviours of using a phone and smartwatch in bed. The fourth factor comprised five items investigated the other behaviours related to the individual's electric light exposure before bedtime. lastly, the fifth factor encompassed six items capturing the individual's morning and daytime light exposure-related behaviour.

@@

While developing and validating LEBA we have complemented the classical test theory based analysis with based analyses (factor analysis and measurement invariance) with IRT analysis. The benefit of implementing IRT analysis was twofold. First, we derived a shorter form of LEBA (18 items). We fitted a graded response model to the combined EFA and CFA sample (n=`r nrow(data)`) and discarded five items (1, 25, 30, 38, & 41) with relatively flat item information curve [I($\theta$) <.20]. The resulting test information curves suggest that the short-LEBA is a psychometrically sound measure with adequate coverage of underlying traits and can be applied to capture the frequency of different light exposure related behaviours reliably. Often, psychological measurements require application of several questionnaires simultaneously. Responding to several lengthy questionnaires increases the participants losing focus and becoming tired. Thus, in some circumstances, reducing the number of items even slightly may be necessary to employ the LEBA questionnaire. Our aim was to provide a data-driven approach to reducing the number of items, apart from the possibility of excluding a specific factor from the 23-item questionnaire. Nonetheless, where possible, we strongly recommend using the extended form of the questionnaire to avoid limiting the range of gained information.

Second, IRT analysis enabled us to acknowledge the individual differences in responding to the items of LEBA. IRT provides a framework to interpret respondents' obtained scores in the light of latent ability (i.e. light exposure behaviour) and the characteristics of the respondents--how they interpret the items [@thomas2019advances]. Findings from the Item and person fit index analysis demonstrate that all five fitted models were acceptable and provide evidence of validity for the factors. In addition, the diverse item discrimination parameters indicate an appropriate range of discrimination -- the ability to differentiate respondents with different levels of light exposure-related behaviour while acknowledging the interpersonal variability in understanding the item.

@@

It is important to note that, LEBA utilizes a five-point Likert-type response scale which is often susceptible to central tendency bias, where responses are biased towards the central value of the response scale [@aston2022central]. Without proper corrections of the response pattern, the interpretation of the scores could be erroneous. We recommend using the sensory precision suggested by @aston2022central method to correct the central tendency bias.

Lastly, this study lacks convergent validity evidence. LEBA being the first of its kind, lacks golden standard against which its convergent validity evidence  could be established. However, a recent study [@Siraji.2023] demonstrates the predictive validity of LEBA by successfully relating LEBA behaviour factors to chronotype, mood, sleep quality, and memory and concentration difficulties. The results of their study confirmed the very essence of LEBA-behaviour could led to different light exposure experiences that differentially influence health, wellness and performance.

@@

To our knowledge, the LEBA is the first inventory characterising light exposure-related behaviour in a scalable manner. Further evidence for the validity of the LEBA could be obtained by administering it conjointly with objective field measurements of light exposure (e.g. with portable light loggers/wearables, see literature review), smartphone readouts as well as subjective data in the form of 24-hour recalls. By this route, one  could study how the (subjectively measured) light exposure-related behavioural patterns translate into (objectively measured) received light exposure, (objectively measured) smartphone use, and how closely the retrospective questionnaire relates to daily reports of these behaviours.

@@

Figure \@ref(fig:FlowchartFig) summarises the steps we followed while developing the LEBA. We conducted all analyses with the statistical software environment R. 
**Firstly**, we set an item pool of 48 items with a six-point Likert-type response format (0-Does not apply/I don't know, 1-Never, 2-Rarely 3-Sometimes, 4-Often, 5-Always) for our initial inventory. Our purpose was to capture light exposure-related behaviour. In that context, the first two response options: "Does not apply/I don't know" and "Never", provided similar information. As such, we collapsed them into one, making it a 5-point Likert-type response format (1-Never, 2-Rarely, 3-Sometimes, 4-Often, 5-Always).

**Secondly**, the two rounds of data collection were administered. In the first round (EFA sample; n=`r nrow(EFA.descriptives)`) we collected data for the exploratory factor analysis (EFA). A sample of at least 250-300 is recommended for EFA [@comreyFirstCourseFactor1992; @schonbrodtWhatSampleSize2013]. The EFA sample exceeded this recommendation. The second round data (CFA sample; n=`r nrow(CFA.descriptives)`) was subjected to confirmatory factor analysis (CFA). To assess sampling adequacy for CFA, we followed the N:q rule [@bentlerPracticalIssuesStructural1987; @jacksonRevisitingSampleSize2003; @klinePrinciplesPracticeStructural2015; @worthingtonScaleDevelopmentResearch2006], where at least ten participants per item are required to earn trustworthiness of the result. Again, our CFA sample exceeded this guidelines.

**Thirdly**, we conducted descriptive and item analyses and proceeded to EFA on the EFA sample. Prior to the EFA, the necessary assumptions, including sample adequacy, normality assumptions, and quality of correlation matrix, were assessed. As our data violated both the univariate and multivariate normality assumption and yielded ordinal response data, we used a polychoric correlation matrix in the EFA and employed "principal axis" (PA) as the factor extraction method [@desjardinsHandbookEducationalMeasurement2018; @watkinsStepbyStepGuideExploratory2020]. We applied a combination of methods, including a Scree plot [@cattellScreeTestNumber1966], minimum average partials method [@velicerDeterminingNumberComponents1976], and Hull method [@lorenzo-sevaHullMethodSelecting2011] to identify factor numbers. To determine the latent structure, we followed the common guidelines: (i) no factors with fewer than three items (ii) no factors with a factor loading \<0.3 (iii) no items with cross-loading \> .3 across factors [@bandalosFactorAnalysisExploratory2018]. <!-- We also conducted an EFA on the non-merged response options data set (**Supplementary File 1**). -->

@@

**Fourthly**, in a secondary analysis, we identified the educational grade level (US education system) required to understand the items in our inventory with the Flesch-Kincaid grade level identification method [@flesch1948new]. Correspondingly, we analysed possible semantic overlap of our developed inventory using the ["Semantic Scale Network" (SSN)](https://rosenbusch.shinyapps.io/semantic_net/) engine [@rosenbusch2020semantic]. The SSN detects semantically related scales and provides a cosine similarity index ranging between -.66 to 1 [@rosenbusch2020semantic]. Pairs of scales with a cosine similarity index value of 1 inidicate full semantical similarity, suggesting redundancy.

**Lastly**, we derived a short form of the LEBA employing an Item Response Theory (IRT) based analysis. We fitted each factor of the LEBA to the combined EFA and CFA sample (n=`r nrow(sem.data)`) using the graded response model [@samejima1997handbook]. IRT assesses the item quality by estimating the item discrimination, item difficulty, item information curve, and test information curve [@bakerBasicsItemResponse2017]. Item discrimination indicates how well a particular item can differentiate between participants across the given latent trait continuum ($\theta$). Item difficulty corresponds to the latent trait level at which the probability of endorsing a particular response option is 50%. The item information curve (IIC) indicates the amount of information an item carries along the latent trait continuum. Here, we reported the item difficulty and discrimination parameter and categorized the items based on their item discrimination index: (i) none=0; (ii) very low=0.01 to 0.34; (iii) low=0.35 to 0.64; (iv) moderate=0.65 to 1.34 ; (v) high=1.35 to 1.69; (vi) very high \>1.70 [@bakerBasicsItemResponse2017]. We discarded the items with a relatively flat item information curve (information \<.2) to derive the short form of LEBA. We also assessed the precision of the short LEBA utilizing the test information curve (TIC). TIC indicates the amount of information a particular scale carries along the latent trait continuum. Additionally, the item and person fit of the fitted IRT models were analysed to gather more evidence on the validity and meaningfulness of our scale [@desjardinsHandbookEducationalMeasurement2018]. The item fit was evaluated using the RMSEA value obtained from Signed-$\chi^2$ index implementation, where an RMSEA value $\le$.06 was considered an adequate item fit. The person fit was estimated employing the standardized fit index Zh statistics [@drasgow1985appropriateness]. Here, Zh \< -2 was considered as a misfit [@drasgow1985appropriateness].

>>>>

#     institution   : "TUM School of Medicine and Health, Technical University of Munich, Munich, Germany"

@@

On the other hand, several attempts have been made to quantify received light exposure subjectively with self-report questionnaires (**Supplementary Table 1**). However, self-reporting light properties could be challenging for people who lack technical knowledge of light sources. Moreover, it is worth considering that the human visual system, unlike a photometer, continuously adapts to ambient brightness [@hurvich_perception_1966], while the signals underlying the non-visual effects of light are independent from perception [@allen_exploiting_2018]. Retrospectively recalling the properties of a light source can further complicate such subjective evaluations. Moreover, measuring light properties alone does not yield any information about how individuals might behave differently regarding diverse light environments such as work, home or outdoors.

To date, little effort has been made to understand and capture these activities. Here, we present the development process of a novel self-reported inventory, the Light Exposure Behaviour Assessment (LEBA), for characterizing diverse light exposure-related behaviours. Notably, the focus of LEBA is not to estimate personal light exposure. Instead, we aim to assess, in a scalable way, how people behave in relation to light, focusing on habitual patterns that could guide behavioural interventions.

@@

Exploratory analysis revealed that items of LEBA inventory can be categorized into five major factors: (i) wearing blue light filters; (ii) spending time out doors; (iii) using phone and smartwatch in bed; (iv) using light before bedtime (v) using light in the morning and during daytime. In this stage of analysis, we retained 25 items. the first factor had three items and encapsulated the individual's preference for using blue light filters in different light environments. The second factor contained six items that incorporated the individuals' hours spent outdoors. The third factor contained five items that looked into specific behaviours of using a phone and smartwatch in bed. The fourth factor comprised five items investigated the other behaviours related to the individual's electric light exposure before bedtime. Lastly, the fifth factor encompassed six items capturing the individual's morning and daytime light exposure-related behaviour.

@@

While developing and validating LEBA, we have complemented conventional approaches with an Item Response Theory (IRT) analysis. IRT provides a framework to interpret respondents' obtained scores in the light of latent ability (i.e. light exposure behaviour) and the characteristics of the respondents -- how they interpret the items [@thomas2019advances]. The benefit of implementing IRT analysis was twofold. First, we derived a shorter form of LEBA (18 items). We fitted a graded response model to the combined EFA and CFA sample (n=`r nrow(data)`) and discarded five items (1, 25, 30, 38, & 41) with relatively flat item information curve [I($\theta$) <.20]. The resulting test information curves suggest that the short-LEBA is a psychometrically sound measure with adequate coverage of underlying traits and can be applied to capture the frequency of different light exposure related behaviours reliably. Often, psychological measurements require application of several questionnaires simultaneously. Responding to several lengthy questionnaires increases the participants losing focus and becoming tired. Thus, in some circumstances, reducing the number of items even slightly may be necessary to employ the LEBA questionnaire. Our aim was to provide a data-driven approach to reducing the number of items, apart from the possibility of excluding a specific factor from the 23-item questionnaire. Nonetheless, where possible, we strongly recommend using the extended form of the questionnaire to avoid limiting the range of gained information.

The IRT analysis enabled us to capture individual differences in responding to the items of LEBA. Findings from the item and person fit index analysis demonstrate that all five fitted models were acceptable and provide evidence of validity for the factors. In addition, the diverse item discrimination parameters indicate an appropriate range of discrimination -- the ability to differentiate respondents with different levels of light exposure-related behaviour while acknowledging the interpersonal variability in understanding the item.

@@

It is important to note that LEBA utilizes a five-point Likert-type response scale which may be susceptible to central tendency bias, i.e. responses are biased towards the central value of the response scale. Future work should evaluate other methods of obtaining responses, such as using a visual-analogue scale.

Finally, there is limited evidence for convergent validity. LEBA being the first of its kind in characterising light exposure _behaviour_ at present lacks a gold standard against which its convergent validity evidence could be established. A recent study [@Siraji.2023] demonstrated the predictive validity of LEBA by successfully relating its factors to self-reported chronotype, mood, sleep quality, and cognitive function. The results of their study confirmed that light-related behaviours, as captured by LEBA, could led to different light exposure experiences that differentially influence health, wellness and performance. Further work will need to establish convergent validity of LEBA further.

@@

To our knowledge, the LEBA is the first inventory characterising light exposure-related behaviour in a scalable manner. Further evidence for the validity of the LEBA could be obtained by administering it conjointly with objective field measurements of light exposure (e.g. with portable light loggers/wearables), smartphone readouts, as well as subjective data in the form of 24-hour recalls. Such a study could related the relationship bewteen (subjectively measured) light exposure-related behavioural patterns translate into (objectively measured) received light exposure, smartphone use, and how closely the retrospective questionnaire relates to daily reports of these behaviours.

@@

Figure \@ref(fig:FlowchartFig) summarises the steps we followed while developing the LEBA. We conducted all analyses with the statistical software environment R.

(1) We set an item pool of 48 items with a six-point Likert-type response format (0-Does not apply/I don't know, 1-Never, 2-Rarely 3-Sometimes, 4-Often, 5-Always) for our initial inventory. Our purpose was to capture light exposure-related behaviour. In that context, the first two response options: "Does not apply/I don't know" and "Never", provided similar information. As such, we collapsed them into one, making it a 5-point Likert-type response format (1-Never, 2-Rarely, 3-Sometimes, 4-Often, 5-Always).

(2) Two rounds of data collection were administered. In the first round (EFA sample; n=`r nrow(EFA.descriptives)`) we collected data for the exploratory factor analysis (EFA). A sample of at least 250-300 is recommended for EFA [@comreyFirstCourseFactor1992; @schonbrodtWhatSampleSize2013]. The EFA sample exceeded this recommendation. The second round data (CFA sample; n=`r nrow(CFA.descriptives)`) was subjected to confirmatory factor analysis (CFA). To assess sampling adequacy for CFA, we followed the N:q rule [@bentlerPracticalIssuesStructural1987; @jacksonRevisitingSampleSize2003; @klinePrinciplesPracticeStructural2015; @worthingtonScaleDevelopmentResearch2006], where at least ten participants per item are required to earn trustworthiness of the result. Again, our CFA sample exceeded this guidelines.

(3) We conducted descriptive and item analyses and proceeded to EFA on the EFA sample. Prior to the EFA, the necessary assumptions, including sample adequacy, normality assumptions, and quality of correlation matrix, were assessed. As our data violated both the univariate and multivariate normality assumption and yielded ordinal response data, we used a polychoric correlation matrix in the EFA and employed "principal axis" (PA) as the factor extraction method [@desjardinsHandbookEducationalMeasurement2018; @watkinsStepbyStepGuideExploratory2020]. We applied a combination of methods, including a Scree plot [@cattellScreeTestNumber1966], minimum average partials method [@velicerDeterminingNumberComponents1976], and Hull method [@lorenzo-sevaHullMethodSelecting2011] to identify factor numbers. To determine the latent structure, we followed the common guidelines: (i) no factors with fewer than three items (ii) no factors with a factor loading \<0.3 (iii) no items with cross-loading \> .3 across factors [@bandalosFactorAnalysisExploratory2018]. <!-- We also conducted an EFA on the non-merged response options data set (**Supplementary File 1**). -->

@@

(4) In a secondary analysis, we identified the educational grade level (US education system) required to understand the items in our inventory with the Flesch-Kincaid grade level identification method [@flesch1948new]. Correspondingly, we analysed possible semantic overlap of our developed inventory using the ["Semantic Scale Network" (SSN)](https://rosenbusch.shinyapps.io/semantic_net/) engine [@rosenbusch2020semantic]. The SSN detects semantically related scales and provides a cosine similarity index ranging between -.66 to 1 [@rosenbusch2020semantic]. Pairs of scales with a cosine similarity index value of 1 inidicate full semantical similarity, suggesting redundancy.

(5) We derived a short form of the LEBA employing an Item Response Theory (IRT) based analysis. We fitted each factor of the LEBA to the combined EFA and CFA sample (n=`r nrow(sem.data)`) using the graded response model [@samejima1997handbook]. IRT assesses the item quality by estimating the item discrimination, item difficulty, item information curve, and test information curve [@bakerBasicsItemResponse2017]. Item discrimination indicates how well a particular item can differentiate between participants across the given latent trait continuum ($\theta$). Item difficulty corresponds to the latent trait level at which the probability of endorsing a particular response option is 50%. The item information curve (IIC) indicates the amount of information an item carries along the latent trait continuum. Here, we reported the item difficulty and discrimination parameter and categorized the items based on their item discrimination index: (i) none=0; (ii) very low=0.01 to 0.34; (iii) low=0.35 to 0.64; (iv) moderate=0.65 to 1.34 ; (v) high=1.35 to 1.69; (vi) very high \>1.70 [@bakerBasicsItemResponse2017]. We discarded the items with a relatively flat item information curve (information \<.2) to derive the short form of LEBA. We also assessed the precision of the short LEBA utilizing the test information curve (TIC). TIC indicates the amount of information a particular scale carries along the latent trait continuum. Additionally, the item and person fit of the fitted IRT models were analysed to gather more evidence on the validity and meaningfulness of our scale [@desjardinsHandbookEducationalMeasurement2018]. The item fit was evaluated using the RMSEA value obtained from Signed-$\chi^2$ index implementation, where an RMSEA value $\le$.06 was considered an adequate item fit. The person fit was estimated employing the standardized fit index Zh statistics [@drasgow1985appropriateness]. Here, Zh \< -2 was considered as a misfit [@drasgow1985appropriateness].
